The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-08 manuscript/manuscript.Rnw
describe n of samples and include t test explanation

\DLA{Add QC information}

@@

For more complete evolutionary comparisons additional sequences of
placenta in \mouse \cite{Wang.ea2011:surveyfornovelimprinted}, \human
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual} \horse
\cite{Coleman.ea2010:Structuralannotationofequine}, \sheep
\cite{Archibald.ea2010:sheepgenomereferencesequence}, and \pig
\cite{Ahanda.ea2012:Predictionofaltered3-} were obtained from the
\ac{SRA} (Supplemental \cref{tab:srr_numbers}).

@@

To determine enrichment levels of functional categories of genes in
highly-expressed transcripts in each transcriptome, we performed
functional annotation cluster analyses per transcriptome using topGO
\cite{topGO:2.20.0}. Gene Ontology clusters were considered
significantly enriched if the \ac{FDR} as estimated by the Benjamini
and Hochberg procedure \cite{Benjamini1995:bh_procedure} on from the
$p$ values obtained from the Fisher Exact Test were less than or equal
to 0.05. Enrichment tests were considered with two gene lists per
transcriptome: annotated genes with expression levels of the top
5\% of the frequency distribution of \ac{FPKM} (Supplemental
\cref{tab:go_terms_enriched_top_one_dasypus_novemcinctus,tab:go_terms_enriched_top_one_equus_caballus,tab:go_terms_enriched_top_one_homo_sapiens,tab:go_terms_enriched_top_one_loxodonta_africana,tab:go_terms_enriched_top_one_monodelphis_domestica,tab:go_terms_enriched_top_one_mus_musculus,tab:go_terms_enriched_top_one_ovis_aries,tab:go_terms_enriched_top_one_pan_paniscus,tab:go_terms_enriched_top_one_sus_scrofa,tab:go_terms_enriched_top_one_ateles_fusciceps,tab:go_terms_enriched_top_one_spalax_galili,tab:go_terms_enriched_top_one_spalax_carmeli,tab:go_terms_enriched_top_one_bos_taurus,tab:go_terms_enriched_top_one_canis_familiaris});
the top 100 genes with the highest \ac{FPKM} (Supplemental
\cref{tab:go_terms_enriched_dasypus_novemcinctus,tab:go_terms_enriched_equus_caballus,tab:go_terms_enriched_homo_sapiens,tab:go_terms_enriched_loxodonta_africana,tab:go_terms_enriched_monodelphis_domestica,tab:go_terms_enriched_mus_musculus,tab:go_terms_enriched_ovis_aries,tab:go_terms_enriched_pan_paniscus,tab:go_terms_enriched_sus_scrofa,tab:go_terms_enriched_ateles_fusciceps,tab:go_terms_enriched_spalax_galili,tab:go_terms_enriched_spalax_carmeli,tab:go_terms_enriched_bos_taurus,tab:go_terms_enriched_canis_familiaris}).
Due to the paucity of annotated gene ontology terms for some species,
gene lists were compared using the \human Gene Ontology
using the 1:1 \human ortholog obtained

@@

>>>>

@@

For more complete evolutionary comparisons additional sequences of
placenta in \mouse
(n=\Sexpr{star.logs[grepl("musculus",species),length(unique(file))]})
\cite{Wang.ea2011:surveyfornovelimprinted}, \human
(n=\Sexpr{star.logs[grepl("sapiens",species),length(unique(file))]})
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual} \horse
\cite{Coleman.ea2010:Structuralannotationofequine}, \sheep
\cite{Archibald.ea2010:sheepgenomereferencesequence}, and \pig
\cite{Ahanda.ea2012:Predictionofaltered3-} were obtained from the
\ac{SRA} (Supplemental \cref{tab:srr_numbers}) and analyzed using the
same methodology as mRNAs sequenced specifically for this study.

@@

To determine enrichment levels of functional categories of genes in
highly-expressed transcripts in each transcriptome, we performed
functional annotation cluster analyses per transcriptome using topGO
\cite{topGO:2.20.0}. Gene Ontology clusters were considered
significantly enriched if the \ac{FDR} as estimated by the Benjamini
and Hochberg procedure \cite{Benjamini1995:bh_procedure} on from the
$p$ values obtained from the Fisher Exact Test were less than or equal
to 0.05. Enrichment tests were considered with two gene lists per
transcriptome: annotated genes with expression levels of the top 5\%
of the frequency distribution of \ac{FPKM} (Supplemental
\cref{tab:go_terms_enriched_top_one_dasypus_novemcinctus,tab:go_terms_enriched_top_one_equus_caballus,tab:go_terms_enriched_top_one_homo_sapiens,tab:go_terms_enriched_top_one_loxodonta_africana,tab:go_terms_enriched_top_one_monodelphis_domestica,tab:go_terms_enriched_top_one_mus_musculus,tab:go_terms_enriched_top_one_ovis_aries,tab:go_terms_enriched_top_one_pan_paniscus,tab:go_terms_enriched_top_one_sus_scrofa,tab:go_terms_enriched_top_one_ateles_fusciceps,tab:go_terms_enriched_top_one_spalax_galili,tab:go_terms_enriched_top_one_spalax_carmeli,tab:go_terms_enriched_top_one_bos_taurus,tab:go_terms_enriched_top_one_canis_familiaris});
the top 100 genes with the highest \ac{FPKM} (Supplemental
\cref{tab:go_terms_enriched_dasypus_novemcinctus,tab:go_terms_enriched_equus_caballus,tab:go_terms_enriched_homo_sapiens,tab:go_terms_enriched_loxodonta_africana,tab:go_terms_enriched_monodelphis_domestica,tab:go_terms_enriched_mus_musculus,tab:go_terms_enriched_ovis_aries,tab:go_terms_enriched_pan_paniscus,tab:go_terms_enriched_sus_scrofa,tab:go_terms_enriched_ateles_fusciceps,tab:go_terms_enriched_spalax_galili,tab:go_terms_enriched_spalax_carmeli,tab:go_terms_enriched_bos_taurus,tab:go_terms_enriched_canis_familiaris}).
Due to the paucity of annotated gene ontology terms for some species,
gene lists were compared using the \human Gene Ontology using the 1:1
\human ortholog. Enrichment was considered out of the set of all 1:1
orthologs present in each species.

@@

Significant shifts in expression levels between groups of species were
calculated using Student's t test on 1:1 orthologs using the mean of
each species to avoid large biases towards samples with $n \gg 1$ and
considered significant if the \ac{FDR} as estimated by the Benjamini
and Hochberg procedure \cite{Benjamini1995:bh_procedure} was
$\le 0.05$.
